---
title: "Sex Differences in Rates of Hearing Decline Before and After Menopause in Koreans"
subtitle: "Comprehensive Statistical Re-analysis"
author: "SungHee Kim et al."
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    df_print: paged
    code_folding: show
  pdf_document:
    toc: true
    toc_depth: 3
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 10,
  fig.height = 6,
  dpi     = 150
)
```

---

## Required Packages

Install if needed:
```r
install.packages(c("tidyverse", "readxl", "MatchIt", "sandwich",
                   "lmtest", "boot", "scales", "patchwork", "writexl"))
```

```{r libraries}
library(tidyverse)
library(readxl)
library(MatchIt)
library(sandwich)
library(lmtest)
library(boot)
library(scales)
library(patchwork)
library(writexl)

# ── Global constants ─────────────────────────────────────────────────────────
BREAKPOINT <- 49.3
FREQS      <- c("0.25kHz", "0.5kHz", "1kHz", "2kHz", "4kHz", "8kHz")
FREQ_LABELS <- c("0.25", "0.5", "1", "2", "4", "8")

COLOR_MEN   <- "#2196F3"
COLOR_WOMEN <- "#E91E63"
SEX_COLORS  <- c("Men" = COLOR_MEN, "Women" = COLOR_WOMEN)
SEX_SHAPES  <- c("Men" = 16, "Women" = 17)

# Clean ggplot2 theme
theme_hearing <- function() {
  theme_classic(base_size = 12) +
    theme(
      strip.background = element_blank(),
      strip.text       = element_text(face = "bold", size = 11),
      legend.position  = "bottom",
      plot.title       = element_text(face = "bold", size = 13),
      plot.subtitle    = element_text(size = 10, color = "grey40"),
      axis.title       = element_text(size = 11)
    )
}

cat("All packages loaded.\n")
```

---

# Phase 1: Data Loading & Cleaning

## 1.1 Load raw data

> **Why:** Before any analysis, we confirm that both files loaded correctly — right number of rows,
> right column names, and no silent data corruption. A single wrong column name here would cause
> every downstream model to fail or silently use the wrong variable.


```{r load-data}
men_raw   <- read_excel("men.xlsx")
women_raw <- read_excel("women.xlsx")

cat("Men shape   :", nrow(men_raw),   "×", ncol(men_raw),   "\n")
cat("Women shape :", nrow(women_raw), "×", ncol(women_raw), "\n")
cat("\nMen columns   :", paste(names(men_raw),   collapse = ", "), "\n")
cat("Women columns :", paste(names(women_raw), collapse = ", "), "\n")
```

## 1.2 Data validation

> **Why:** We already know from prior inspection that `women.xlsx` previously contained a
> duplicate block (rows 3,180–6,354 were an exact copy of rows 1–3,175 shifted one column right).
> Dr. Kim removed this manually, so this cell confirms the fix held and that neither file has
> unexpected nulls or duplicate rows before we proceed.


```{r validate}
validate_dataset <- function(df, name) {
  cat("===", name, "===\n")
  cat("  Shape    :", nrow(df), "×", ncol(df), "\n")
  cat("  Columns  :", paste(names(df), collapse = ", "), "\n")

  missing_freq <- setdiff(FREQS, names(df))
  if (length(missing_freq) > 0)
    cat("  ⚠ Missing frequency columns:", paste(missing_freq, collapse = ", "), "\n")

  null_counts <- colSums(is.na(df))
  if (any(null_counts > 0))
    cat("  Nulls    :", paste(names(null_counts[null_counts > 0]),
                              null_counts[null_counts > 0], sep = "=", collapse = ", "), "\n")
  else
    cat("  No missing values\n")

  n_dup <- sum(duplicated(df))
  cat("  Duplicates:", n_dup, "\n")
  cat("  Age range :", round(min(df$Age), 1), "–", round(max(df$Age), 1), "yr\n\n")
}

validate_dataset(men_raw,   "men.xlsx")
validate_dataset(women_raw, "women.xlsx")
```

## 1.3 Standardise and merge

> **How it works:** We assign a `Sex` label to each dataset, select only the columns we need
> (`Age` and the 6 frequency thresholds), and stack them into one unified dataframe. We also
> create two helper columns: `SexCode` (0 = Men, 1 = Women) for regression use, and `AgeGroup`
> to flag participants as being below or above the 49.3-year menopause breakpoint.


```{r merge}
standardise_df <- function(df, sex_label) {
  df %>%
    rename_with(str_trim) %>%
    filter(!is.na(Age)) %>%
    select(all_of(c("Age", FREQS))) %>%
    mutate(Sex = sex_label)
}

men_df   <- standardise_df(men_raw,   "Men")
women_df <- standardise_df(women_raw, "Women")

df <- bind_rows(men_df, women_df) %>%
  mutate(
    Sex      = factor(Sex, levels = c("Men", "Women")),
    SexCode  = as.integer(Sex == "Women"),
    AgeGroup = if_else(Age < BREAKPOINT,
                       paste0("< ", BREAKPOINT, " yr"),
                       paste0(">= ", BREAKPOINT, " yr")),
    AgeGroup = factor(AgeGroup,
                      levels = c(paste0("< ", BREAKPOINT, " yr"),
                                 paste0(">= ", BREAKPOINT, " yr"))),
    Decade   = factor(paste0(floor(Age / 10) * 10, "s"))
  )

df %>% count(Sex) %>% rename(N = n)
```

## 1.4 Basic data quality summary

```{r quality}
df %>%
  group_by(Sex) %>%
  summarise(
    N    = n(),
    Age_mean = round(mean(Age), 1),
    Age_sd   = round(sd(Age), 1),
    Age_min  = round(min(Age), 1),
    Age_max  = round(max(Age), 1),
    .groups = "drop"
  )
```

```{r threshold-summary}
df %>%
  pivot_longer(all_of(FREQS), names_to = "Freq", values_to = "Threshold") %>%
  group_by(Sex, Freq) %>%
  summarise(
    Mean = round(mean(Threshold, na.rm = TRUE), 2),
    SD   = round(sd(Threshold, na.rm = TRUE), 2),
    Min  = round(min(Threshold, na.rm = TRUE), 2),
    Max  = round(max(Threshold, na.rm = TRUE), 2),
    .groups = "drop"
  )
```

---

# Phase 2: Exploratory Data Analysis

## 2.1 Age distribution by sex

> **Why:** Men (n = 910) and Women (n = 3,175) are unequal in size — roughly 1:3.5. Before
> modelling, we check whether this imbalance also means the two groups have different age
> *distributions*. If women are systematically younger, any hearing difference we observe could
> partly reflect age rather than sex. The dashed vertical line marks the 49.3-year menopause
> breakpoint, so we can also see how many participants fall on each side.
>
> **Interpretation:** Look for whether the two histograms have similar shapes and centres.
> Substantial differences (e.g. women skewed toward younger ages) would justify PSM in Phase 3.


```{r fig-age-dist, fig.height=4}
ggplot(df, aes(x = Age, fill = Sex)) +
  geom_histogram(bins = 30, alpha = 0.75, color = "white") +
  geom_vline(xintercept = BREAKPOINT, linetype = "dashed",
             color = "black", linewidth = 0.8) +
  annotate("text", x = BREAKPOINT + 1, y = Inf,
           label = paste0("Breakpoint\n(", BREAKPOINT, " yr)"),
           vjust = 1.5, hjust = 0, size = 3.5) +
  facet_wrap(~ Sex, scales = "free_y") +
  scale_fill_manual(values = SEX_COLORS) +
  labs(title = "Age Distribution by Sex",
       x = "Age (years)", y = "Count") +
  theme_hearing() +
  theme(legend.position = "none")
```

## 2.2 Figure 2 — Audiogram: Mean hearing threshold by age decade

> **Why:** The audiogram is the standard clinical display of hearing data. Plotting mean
> thresholds by age decade lets us see the trajectory of age-related hearing loss across the full
> frequency range, separately for men and women. This replicates Figure 2 in the manuscript
> (Fig 2 audiogram.pptx).
>
> **How to read an audiogram:** The y-axis is *inverted* — better hearing (lower dB) is at the
> top, worse hearing (higher dB) is at the bottom. Each line represents one decade of life.
> Lines moving downward as age increases = hearing getting worse.
>
> **What to look for:** High frequencies (4 kHz, 8 kHz) typically show the steepest decline
> with age. If the women's panel shows a noticeable shift in the line spacing around the 50s
> decade, that is consistent with the menopause hypothesis.
 — Audiogram: Mean hearing threshold by age decade

*Replicates Fig 2 audiogram.pptx — audiogram convention: y-axis inverted (worse hearing at bottom)*

```{r fig2-audiogram, fig.height=9, fig.width=10}
# Compute mean ± SD by sex, decade, frequency
audiogram_data <- df %>%
  pivot_longer(all_of(FREQS), names_to = "Freq", values_to = "Threshold") %>%
  mutate(Freq = factor(Freq, levels = FREQS)) %>%
  group_by(Sex, Decade, Freq) %>%
  summarise(
    Mean = mean(Threshold, na.rm = TRUE),
    SD   = sd(Threshold, na.rm = TRUE),
    N    = n(),
    SE   = SD / sqrt(N),
    .groups = "drop"
  )

# Decade colour palette
decade_levels <- levels(df$Decade)
decade_colors <- colorRampPalette(c("#4FC3F7", "#0D47A1"))(length(decade_levels))
names(decade_colors) <- decade_levels

make_audiogram <- function(sex_label) {
  audiogram_data %>%
    filter(Sex == sex_label) %>%
    ggplot(aes(x = Freq, y = Mean, color = Decade, group = Decade)) +
    geom_line(linewidth = 0.9) +
    geom_point(size = 2.5) +
    geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                  width = 0.2, alpha = 0.6) +
    scale_y_reverse(
      limits = c(80, -10),
      breaks = seq(-10, 80, by = 10),
      labels = seq(-10, 80, by = 10)
    ) +
    scale_color_manual(values = decade_colors) +
    labs(
      title    = paste0("(", ifelse(sex_label == "Men", "A", "B"), ") ", sex_label),
      x        = "Frequency (kHz)",
      y        = "Hearing Level (dB HL)",
      color    = "Age decade"
    ) +
    theme_hearing() +
    theme(legend.position = "right")
}

fig2_men   <- make_audiogram("Men")
fig2_women <- make_audiogram("Women")

fig2_men / fig2_women +
  plot_annotation(
    title    = "Figure 2. Mean Hearing Thresholds by Age Decade",
    subtitle = "Mean ± SD; y-axis inverted per audiogram convention (worse hearing at bottom)",
    theme    = theme(plot.title = element_text(face = "bold", size = 13))
  )
```

## 2.3 Hearing threshold distributions by frequency

> **Why:** Before fitting any regression model, we inspect the raw shape of the outcome variable
> at each frequency. OLS regression assumes the residuals (not the raw data, but close) are
> approximately normally distributed. Right-skewed or heavy-tailed distributions are a warning
> sign that the standard errors and p-values from plain OLS may be unreliable.
>
> **Interpretation:** A roughly bell-shaped, symmetric histogram at each frequency is ideal.
> A long right tail (most values cluster low, but some extend far right) is common in audiological
> data and indicates positive skewness — which we address in Phase 5 with WLS and robust SEs.


```{r fig-distributions, fig.height=7}
df %>%
  pivot_longer(all_of(FREQS), names_to = "Freq", values_to = "Threshold") %>%
  mutate(Freq = factor(Freq, levels = FREQS)) %>%
  ggplot(aes(x = Threshold, fill = Sex)) +
  geom_histogram(aes(y = after_stat(density)),
                 bins = 40, alpha = 0.55, color = "white",
                 position = "identity") +
  facet_wrap(~ Freq, nrow = 2) +
  scale_fill_manual(values = SEX_COLORS) +
  labs(title = "Hearing Threshold Distributions by Frequency and Sex",
       x = "Threshold (dB HL)", y = "Density") +
  theme_hearing()
```

## 2.4 Skewness and kurtosis

> **Why:** Skewness and kurtosis give us numerical summaries of distributional shape. OLS is
> robust to mild non-normality when sample sizes are large, but severe skewness (> 1.0) or excess
> kurtosis (> 3.0) in the residuals can inflate Type I error rates and produce misleading
> confidence intervals.
>
> **How to interpret:**
>
> - **Skewness = 0** → perfectly symmetric. Positive skewness means a right tail (some participants
>   have very high thresholds). Values above 1.0 are considered moderately skewed.
> - **Excess kurtosis = 0** → normal-like tails. Positive values mean heavier tails than a normal
>   distribution (more extreme values than expected). Values above 3.0 are a concern.
>
> These numbers directly motivate the remedies applied in Phase 5.


```{r skew-kurt}
skew_kurt <- df %>%
  pivot_longer(all_of(FREQS), names_to = "Freq", values_to = "Threshold") %>%
  group_by(Freq) %>%
  summarise(
    Skewness        = round(e1071::skewness(Threshold, na.rm = TRUE), 3),
    Excess_Kurtosis = round(e1071::kurtosis(Threshold, na.rm = TRUE), 3),
    .groups = "drop"
  )

# Install e1071 if needed: install.packages("e1071")
# Alternative without e1071:
skew_kurt <- df %>%
  pivot_longer(all_of(FREQS), names_to = "Freq", values_to = "Threshold") %>%
  group_by(Freq) %>%
  summarise(
    Skewness = round({
      x <- na.omit(Threshold)
      n <- length(x)
      m <- mean(x); s <- sd(x)
      (n / ((n-1)*(n-2))) * sum(((x - m)/s)^3)
    }, 3),
    Excess_Kurtosis = round({
      x <- na.omit(Threshold)
      n <- length(x)
      m <- mean(x); s <- sd(x)
      ((n*(n+1))/((n-1)*(n-2)*(n-3))) * sum(((x-m)/s)^4) -
        (3*(n-1)^2)/((n-2)*(n-3))
    }, 3),
    .groups = "drop"
  )

print(skew_kurt)

skew_kurt %>%
  pivot_longer(-Freq, names_to = "Metric", values_to = "Value") %>%
  mutate(Freq = factor(Freq, levels = FREQS)) %>%
  ggplot(aes(x = Freq, y = Value, fill = Metric)) +
  geom_col(position = "dodge", alpha = 0.85) +
  geom_hline(yintercept = 0, linewidth = 0.5) +
  scale_fill_manual(values = c("Skewness" = "steelblue",
                               "Excess_Kurtosis" = "coral")) +
  labs(title = "Skewness and Excess Kurtosis by Frequency",
       x = "Frequency", y = "Value", fill = "") +
  theme_hearing()
```

## 2.5 Correlation heatmap

> **Why:** We check whether age is meaningfully correlated with hearing thresholds at each
> frequency — if the correlation is near zero, a regression of threshold on age would be
> uninformative. We also check whether the six frequencies are highly correlated with each other,
> which would be relevant if we were fitting a multivariate model.
>
> **Interpretation:** Values close to +1 (dark red) indicate a strong positive relationship —
> older age associated with higher (worse) hearing thresholds. Values near 0 indicate no linear
> association. We expect the highest correlations at 4 kHz and 8 kHz, where age-related decline
> is typically fastest.


```{r fig-corr, fig.height=5}
# Requires reshape2 or use base R
make_corr_plot <- function(sex_label) {
  corr_data <- df %>%
    filter(Sex == sex_label) %>%
    select(Age, all_of(FREQS)) %>%
    cor(use = "complete.obs") %>%
    as.data.frame() %>%
    rownames_to_column("Var1") %>%
    pivot_longer(-Var1, names_to = "Var2", values_to = "Corr") %>%
    mutate(
      Var1 = factor(Var1, levels = c("Age", FREQS)),
      Var2 = factor(Var2, levels = c("Age", FREQS))
    )

  ggplot(corr_data, aes(x = Var1, y = Var2, fill = Corr)) +
    geom_tile(color = "white") +
    geom_text(aes(label = round(Corr, 2)), size = 3) +
    scale_fill_gradient2(low = "#2166AC", mid = "white", high = "#D6604D",
                         midpoint = 0, limits = c(-1, 1)) +
    labs(title = sex_label, x = NULL, y = NULL, fill = "r") +
    theme_hearing() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "right")
}

make_corr_plot("Men") + make_corr_plot("Women") +
  plot_annotation(title = "Pearson Correlation: Age vs Hearing Thresholds",
                  theme = theme(plot.title = element_text(face = "bold")))
```

## 2.6 Mean threshold by age decade

```{r fig-by-decade, fig.height=7}
decade_means <- df %>%
  pivot_longer(all_of(FREQS), names_to = "Freq", values_to = "Threshold") %>%
  mutate(Freq = factor(Freq, levels = FREQS)) %>%
  group_by(Sex, Decade, Freq) %>%
  summarise(Mean = mean(Threshold, na.rm = TRUE), .groups = "drop")

ggplot(decade_means, aes(x = Decade, y = Mean, color = Sex, group = Sex)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  facet_wrap(~ Freq, nrow = 2) +
  scale_color_manual(values = SEX_COLORS) +
  labs(title = "Mean Hearing Threshold by Age Decade and Frequency",
       x = "Age Decade", y = "Mean Threshold (dB HL)") +
  theme_hearing()
```

---

# Phase 3: Imbalance Handling — Propensity Score Matching (PSM)

## 3.1 Estimate propensity scores and match

> **Why this method was chosen:** With 910 men and 3,175 women, the groups are imbalanced ~1:3.5.
> If the age distributions also differ between groups, a direct sex comparison confounds sex with
> age. Propensity Score Matching (PSM) creates a balanced subsample where men and women are
> virtually identical in age, removing this confound.
>
> **How it works:**
>
> 1. For each participant, a logistic regression estimates the probability of being female given
>    their age: P(female | Age). This is the **propensity score**.
> 2. Each man is then matched to the woman with the closest propensity score (nearest-neighbour
>    matching), subject to a **caliper** — a maximum allowed distance of 0.2 standard deviations
>    of the logit-transformed propensity score. Matches outside this caliper are discarded.
> 3. This produces a 1:1 matched sample where the two groups are age-balanced.
>
> **Important:** PSM is used only as a **sensitivity analysis**. The primary analysis runs on the
> full unmatched sample (n = 4,085). If the matched and unmatched results agree, it confirms that
> the sex imbalance was not driving the findings.


```{r psm}
# PSM using MatchIt: 1:1 nearest-neighbour on Age
psm_out <- matchit(
  SexCode ~ Age,
  data    = df,
  method  = "nearest",
  ratio   = 1,
  caliper = 0.2,       # 0.2 × SD of logit(PS)
  std.caliper = TRUE
)

summary(psm_out)

df_matched <- match.data(psm_out)
cat("\nMatched sample sizes:\n")
print(table(df_matched$Sex))
```

## 3.2 Check balance — standardised mean difference (SMD)

> **How to interpret SMD:** The Standardised Mean Difference measures how far apart the group
> means are in units of the pooled standard deviation — like a Cohen's d for the covariate.
>
> - **SMD < 0.1** → well-balanced (the groups are effectively equivalent on that variable)
> - **SMD 0.1–0.2** → mild imbalance
> - **SMD > 0.2** → meaningful imbalance that could confound comparisons
>
> After successful PSM, the SMD for Age should drop well below 0.1, confirming the two groups
> are now comparable.
 — standardised mean difference (SMD)

```{r psm-balance}
compute_smd <- function(data, var = "Age") {
  men   <- data[[var]][data$Sex == "Men"]
  women <- data[[var]][data$Sex == "Women"]
  pooled_sd <- sqrt((var(men) + var(women)) / 2)
  abs(mean(men) - mean(women)) / pooled_sd
}

cat("SMD (Age) before PSM:", round(compute_smd(df), 3),
    " — (< 0.1 = well-balanced)\n")
cat("SMD (Age) after  PSM:", round(compute_smd(df_matched), 3), "\n")
```

```{r fig-psm-balance, fig.height=4}
bind_rows(
  df         %>% mutate(Sample = "Before PSM"),
  df_matched %>% mutate(Sample = "After PSM")
) %>%
  mutate(Sample = factor(Sample, levels = c("Before PSM", "After PSM"))) %>%
  ggplot(aes(x = Age, fill = Sex)) +
  geom_histogram(aes(y = after_stat(density)),
                 bins = 30, alpha = 0.6, color = "white",
                 position = "identity") +
  facet_grid(Sex ~ Sample) +
  scale_fill_manual(values = SEX_COLORS) +
  labs(title = "Age Distribution Before vs. After PSM",
       x = "Age (years)", y = "Density") +
  theme_hearing() +
  theme(legend.position = "none")
```

---

# Phase 4: Regression Assumption Checking

## 4.1 Helper functions: Piecewise linear regression setup

> **Why piecewise linear regression?**
>
> The central hypothesis of this study is that the *rate* of hearing decline changes at the age
> of menopause (49.3 years). A single straight line across all ages would force a constant rate
> and miss this inflection. Piecewise LR solves this by fitting two separate line segments — one
> for ages below 49.3 yr, one for ages above — joined at the breakpoint.
>
> **How it works mathematically:**
>
> The model is: `Threshold = B0 + B1 × Age_below + B2 × Age_above`
>
> Where:
>
> - `Age_below = min(Age, 49.3)` — increases up to the breakpoint, then stays flat
> - `Age_above = max(Age − 49.3, 0)` — is 0 below the breakpoint, then increases linearly
>
> So **B1** is the slope (dB/year) *before* menopause, and **B2** is the slope *after* menopause.
> This is mathematically identical to splitting the dataset at 49.3 and running two separate
> regressions — which is exactly what Dr. Kim did in the original manuscript.
>
> **The breakpoint (49.3 yr)** is fixed a priori based on the established Korean mean menopause
> age (Park et al., 2017), not estimated from the data. This is appropriate for a
> population-level study and avoids overfitting.


```{r helpers}
# Build piecewise design matrix
make_piecewise_df <- function(data, freq, bp = BREAKPOINT) {
  data %>%
    select(Age, all_of(freq)) %>%
    rename(Threshold = all_of(freq)) %>%
    filter(!is.na(Threshold)) %>%
    mutate(
      Age_below = pmin(Age, bp),
      Age_above = pmax(Age - bp, 0)
    )
}

# Fit piecewise OLS
fit_piecewise_ols <- function(data, freq, bp = BREAKPOINT) {
  d <- make_piecewise_df(data, freq, bp)
  lm(Threshold ~ Age_below + Age_above, data = d)
}

# Fit piecewise OLS with HC3 robust SEs
fit_piecewise_robust <- function(data, freq, bp = BREAKPOINT) {
  mod <- fit_piecewise_ols(data, freq, bp)
  coeftest(mod, vcov = vcovHC(mod, type = "HC3"))
}

# Fit piecewise WLS (inverse-variance weights by 5-yr age bin)
fit_piecewise_wls <- function(data, freq, bp = BREAKPOINT) {
  d <- make_piecewise_df(data, freq, bp)
  d <- d %>%
    mutate(bin = cut(Age, breaks = seq(15, 85, 5))) %>%
    group_by(bin) %>%
    mutate(bin_var = var(Threshold, na.rm = TRUE)) %>%
    ungroup() %>%
    mutate(weight = 1 / pmax(bin_var, 1e-4))
  lm(Threshold ~ Age_below + Age_above, data = d, weights = d$weight)
}

cat("Helper functions defined.\n")
```

## 4.2 Normality of residuals — Shapiro-Wilk test

> **Why we check this:** OLS regression relies on the assumption that residuals (the difference
> between observed and model-predicted values) are approximately normally distributed. If they
> are not, the p-values and confidence intervals produced by standard OLS may be wrong —
> particularly in smaller samples.
>
> **How the test works:** The Shapiro-Wilk test compares the distribution of residuals to a
> theoretical normal distribution. A small p-value (< 0.05) means the residuals are
> significantly non-normal.
>
> **Interpretation:**
>
> - p > 0.05 → residuals are consistent with normality → standard OLS inference is safe
> - p < 0.05 → evidence of non-normality → bootstrap CIs and robust SEs (Phase 5) become
>   especially important as they do not assume normality
>
> Note: With large samples (Women n = 3,175), the Shapiro-Wilk test becomes very sensitive and
> will almost always reject normality even for mild deviations. The Q-Q plots (Section 4.3)
> give a more practical visual assessment.
 — Shapiro-Wilk test

```{r normality}
normality_results <- map_dfr(c("Men", "Women"), function(sex) {
  map_dfr(FREQS, function(freq) {
    mod   <- fit_piecewise_ols(df %>% filter(Sex == sex), freq)
    resid <- residuals(mod)
    if (length(resid) <= 5000) {
      test_res <- shapiro.test(resid)
      tibble(Sex = sex, Freq = freq, Test = "Shapiro-Wilk",
             Statistic = round(test_res$statistic, 4),
             p_value   = round(test_res$p.value, 4))
    } else {
      test_res <- ks.test(scale(resid), "pnorm")
      tibble(Sex = sex, Freq = freq, Test = "KS",
             Statistic = round(test_res$statistic, 4),
             p_value   = round(test_res$p.value, 4))
    }
  })
}) %>%
  mutate(Normal = if_else(p_value > 0.05, "Yes", "No"))

print(normality_results)
```

## 4.3 QQ-plots of residuals

> **How to read a Q-Q plot:** Each point represents one residual. The x-axis is where that
> residual *would* fall if the data were perfectly normal; the y-axis is where it *actually*
> falls. If the residuals are normal, all points lie on the diagonal reference line.
>
> **Patterns to look for:**
>
> - **Points hugging the line** → normality assumption holds
> - **S-shaped curve** → skewness (residuals have a long tail on one side)
> - **Points curving away at both ends** → heavy tails / excess kurtosis
> - **A few extreme points at the ends** → outliers
>
> Hearing threshold data typically shows S-shaped or upward-curving Q-Q plots due to the
> positive skewness seen in Section 2.4.
 of residuals

```{r fig-qq, fig.height=7, fig.width=12}
qq_data <- map_dfr(c("Men", "Women"), function(sex) {
  map_dfr(FREQS, function(freq) {
    mod   <- fit_piecewise_ols(df %>% filter(Sex == sex), freq)
    resid <- scale(residuals(mod))
    n     <- length(resid)
    tibble(
      Sex       = sex,
      Freq      = freq,
      Observed  = sort(resid),
      Theoretical = qnorm(ppoints(n))
    )
  })
}) %>%
  mutate(Freq = factor(Freq, levels = FREQS),
         Sex  = factor(Sex, levels = c("Men", "Women")))

ggplot(qq_data, aes(x = Theoretical, y = Observed, color = Sex)) +
  geom_point(alpha = 0.3, size = 0.8) +
  geom_abline(slope = 1, intercept = 0, color = "black", linewidth = 0.6) +
  facet_grid(Sex ~ Freq) +
  scale_color_manual(values = SEX_COLORS) +
  labs(title = "Q-Q Plots of OLS Residuals (Piecewise LR)",
       x = "Theoretical Quantiles", y = "Standardised Residuals") +
  theme_hearing() +
  theme(legend.position = "none")
```

## 4.4 Homoscedasticity — Breusch-Pagan test

> **Why we check this:** OLS assumes that residual variance is *constant* across all fitted
> values (homoscedasticity). In hearing data, variance typically grows with age — younger
> people cluster tightly around similar low thresholds, while older people show much wider
> spread. This is called **heteroscedasticity**, and it makes standard OLS standard errors
> too small (overconfident), inflating the number of significant results.
>
> **How the Breusch-Pagan test works:** It regresses the squared residuals on the predictor
> variables. If the predictors explain the squared residuals well (large test statistic, small
> p-value), the variance is not constant — heteroscedasticity is present.
>
> **Interpretation:**
>
> - p > 0.05 → no evidence of heteroscedasticity → standard OLS SEs are reliable
> - p < 0.05 → heteroscedasticity detected → use WLS or HC3 robust SEs (Phase 5)
 — Breusch-Pagan test

```{r breusch-pagan}
# Requires lmtest
hetero_results <- map_dfr(c("Men", "Women"), function(sex) {
  map_dfr(FREQS, function(freq) {
    mod <- fit_piecewise_ols(df %>% filter(Sex == sex), freq)
    bp  <- lmtest::bptest(mod)
    tibble(
      Sex            = sex,
      Freq           = freq,
      BP_statistic   = round(bp$statistic, 3),
      BP_p_value     = round(bp$p.value, 4),
      Heteroscedastic = if_else(bp$p.value < 0.05, "Yes", "No")
    )
  })
})

print(hetero_results)
```

## 4.5 Residual vs. fitted plots

> **How to read these plots:** The x-axis shows the model's predicted hearing threshold;
> the y-axis shows how far each observed value is from that prediction (the residual).
> If the model is well-specified, the residuals should scatter randomly around the horizontal
> zero line with roughly equal spread everywhere.
>
> **Warning signs:**
>
> - **Fan shape** (spread increases as fitted values increase) → heteroscedasticity
> - **Curve or arch** → the true relationship is non-linear and the model is missing it
> - **A few extreme points far from zero** → influential outliers
 plots

```{r fig-resid-fitted, fig.height=7, fig.width=12}
resid_data <- map_dfr(c("Men", "Women"), function(sex) {
  map_dfr(FREQS, function(freq) {
    mod <- fit_piecewise_ols(df %>% filter(Sex == sex), freq)
    tibble(
      Sex     = sex,
      Freq    = freq,
      Fitted  = fitted(mod),
      Residual = residuals(mod)
    )
  })
}) %>%
  mutate(Freq = factor(Freq, levels = FREQS),
         Sex  = factor(Sex, levels = c("Men", "Women")))

ggplot(resid_data, aes(x = Fitted, y = Residual, color = Sex)) +
  geom_point(alpha = 0.25, size = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.6) +
  facet_grid(Sex ~ Freq) +
  scale_color_manual(values = SEX_COLORS) +
  labs(title = "Residual vs. Fitted (Piecewise LR)",
       x = "Fitted Values", y = "Residuals") +
  theme_hearing() +
  theme(legend.position = "none")
```

## 4.6 Residual SD by age decade

```{r fig-hetero-decade, fig.height=7}
sd_decade <- df %>%
  pivot_longer(all_of(FREQS), names_to = "Freq", values_to = "Threshold") %>%
  mutate(
    Freq   = factor(Freq, levels = FREQS),
    Decade = floor(Age / 10) * 10
  ) %>%
  group_by(Sex, Freq, Decade) %>%
  summarise(SD = sd(Threshold, na.rm = TRUE), .groups = "drop")

ggplot(sd_decade, aes(x = Decade, y = SD, color = Sex, group = Sex)) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  facet_wrap(~ Freq, nrow = 2) +
  scale_color_manual(values = SEX_COLORS) +
  labs(title = "SD of Hearing Thresholds by Age Decade (Heteroscedasticity)",
       x = "Age Decade", y = "SD (dB HL)") +
  theme_hearing()
```

## 4.7 Outlier detection — Cook's Distance

> **Why Cook's Distance:** A simple threshold on residual size catches large errors but misses
> *influential* points — observations that, if removed, would substantially change the
> regression slope. Cook's Distance combines residual size with leverage (how unusual the
> predictor value is) into a single influence measure.
>
> **How to interpret:**
>
> - The conventional threshold is **4/n** (where n = sample size). Points above this are
>   flagged as potentially influential.
> - A small percentage of outliers (< 5%) in a large dataset is generally tolerable, especially
>   if results are confirmed with robust methods.
> - High Cook's D values at specific age–frequency combinations (e.g. elderly men at 8 kHz)
>   may reflect genuine clinical outliers rather than data errors.
 — Cook's Distance

```{r cooks-d}
cooks_results <- map_dfr(c("Men", "Women"), function(sex) {
  map_dfr(FREQS, function(freq) {
    mod       <- fit_piecewise_ols(df %>% filter(Sex == sex), freq)
    cooks     <- cooks.distance(mod)
    threshold <- 4 / length(cooks)
    n_out     <- sum(cooks > threshold)
    tibble(
      Sex        = sex,
      Freq       = freq,
      N_outliers = n_out,
      Threshold  = round(threshold, 5),
      Pct        = round(100 * n_out / length(cooks), 2)
    )
  })
})

print(cooks_results)
```

---

# Phase 5: Violation Remedies

## 5.1 Bootstrap percentile confidence intervals for slopes

> **Why bootstrap instead of standard CIs?**
>
> Standard OLS confidence intervals assume residuals are normally distributed. We have already
> seen that this assumption is violated (positive skewness, excess kurtosis). Bootstrap CIs
> make **no distributional assumptions** — they estimate the sampling variability of the slope
> empirically by repeatedly resampling the data.
>
> **How bootstrap works:**
>
> 1. Randomly draw n observations *with replacement* from the dataset (a "bootstrap sample").
> 2. Fit the piecewise regression on this resample and record the slope.
> 3. Repeat 2,000 times.
> 4. The 2.5th and 97.5th percentiles of the 2,000 slope estimates form the 95% confidence
>    interval.
>
> The `cache=TRUE` option means RStudio saves the bootstrap results after the first run, so
> re-knitting the document does not repeat the 2,000 iterations.
>
> **Interpretation:** A bootstrap CI for a slope that does not include zero means the rate of
> decline is statistically significantly different from zero at the 5% level — without relying
> on the normality assumption.
 confidence intervals for slopes

```{r bootstrap, cache=TRUE}
bootstrap_slopes <- function(data, freq, bp = BREAKPOINT,
                              n_boot = 2000, seed = 42) {
  set.seed(seed)
  d <- make_piecewise_df(data, freq, bp)

  get_slopes <- function(data, indices) {
    d_boot <- data[indices, ]
    mod    <- lm(Threshold ~ Age_below + Age_above, data = d_boot)
    coef(mod)[c("Age_below", "Age_above")]
  }

  boot_res <- boot(d, get_slopes, R = n_boot)

  ci_below <- boot.ci(boot_res, type = "perc", index = 1)$percent[4:5]
  ci_above <- boot.ci(boot_res, type = "perc", index = 2)$percent[4:5]

  list(
    slope_below = coef(lm(Threshold ~ Age_below + Age_above, data = d))["Age_below"],
    slope_above = coef(lm(Threshold ~ Age_below + Age_above, data = d))["Age_above"],
    CI_below    = ci_below,
    CI_above    = ci_above,
    boot_obj    = boot_res
  )
}

cat("Running bootstrap (2000 resamples × 12 combinations — ~1-2 min)...\n")

boot_results <- list()
for (sex in c("Men", "Women")) {
  for (freq in FREQS) {
    key <- paste(sex, freq, sep = "_")
    boot_results[[key]] <- bootstrap_slopes(
      df %>% filter(Sex == sex), freq, n_boot = 2000
    )
    r <- boot_results[[key]]
    cat(sprintf("%-6s %-8s | Below: %+.3f [%+.3f, %+.3f]  Above: %+.3f [%+.3f, %+.3f]\n",
                sex, freq,
                r$slope_below, r$CI_below[1], r$CI_below[2],
                r$slope_above, r$CI_above[1], r$CI_above[2]))
  }
}
```

## 5.2 FDR correction — Benjamini-Hochberg (BH) procedure

> **Why this is necessary:** We run ANCOVA separately for each of the 6 frequencies × 2 age
> groups = 12 tests. At α = 0.05, running 12 independent tests gives a ~46% chance of finding
> at least one false positive by chance alone (family-wise error). Simply reporting each p-value
> separately without correction inflates the false discovery rate.
>
> **How Benjamini-Hochberg works:**
>
> 1. Rank all p-values from smallest to largest.
> 2. Each p-value is compared to a threshold of (rank / total tests) × 0.05.
> 3. The largest p-value that is still below its threshold is declared significant, and all
>    smaller p-values are also declared significant.
>
> BH controls the **False Discovery Rate (FDR)** — the expected proportion of significant
> results that are actually false — rather than the probability of any false positive (FWER).
> It is less conservative than Bonferroni correction and is the standard approach in
> exploratory biomedical research.
>
> **Interpretation:** Focus on `FDR_p` rather than `Raw_p`. A result with Raw_p = 0.02 but
> FDR_p = 0.08 should *not* be reported as significant after correcting for multiple comparisons.
 — ANCOVA sex differences

```{r ancova-fdr}
collect_ancova_pvalues <- function(data, bp = BREAKPOINT) {
  map_dfr(list(c("below", paste0("< ", bp)), c("above", paste0(">= ", bp))),
          function(grp) {
    sub <- if (grp[1] == "below") filter(data, Age < bp) else filter(data, Age >= bp)
    map_dfr(FREQS, function(freq) {
      d   <- sub %>% select(SexCode, Age, all_of(freq)) %>%
               rename(Threshold = all_of(freq)) %>% drop_na()
      mod <- lm(Threshold ~ SexCode + Age, data = d)
      rob <- coeftest(mod, vcov = vcovHC(mod, type = "HC3"))
      tibble(
        AgeGroup     = grp[2],
        Freq         = freq,
        SexCoef_dB   = round(rob["SexCode", "Estimate"], 3),
        Raw_p        = round(rob["SexCode", "Pr(>|t|)"], 4)
      )
    })
  })
}

ancova_df <- collect_ancova_pvalues(df) %>%
  mutate(
    FDR_p       = round(p.adjust(Raw_p, method = "BH"), 4),
    Significant = FDR_p < 0.05
  )

cat("=== ANCOVA: Sex Differences (HC3 SEs, BH FDR-corrected) ===\n")
print(ancova_df)
```

---

# Phase 6: Model Building & Visualisation

## 6.1 Figure 3 — Piecewise LR scatterplots

> **What this figure shows:** For each of the 6 audiometric frequencies, individual hearing
> thresholds (dots) are plotted against age, with two separate regression lines per sex — one
> for ages below 49.3 yr (solid segment) and one for ages above (continuing segment). The
> vertical dotted line marks the 49.3-year breakpoint.
>
> **Why this method was chosen:** A single regression line across all ages would miss the
> change in slope at menopause — the key hypothesis of this paper. Two separate segments,
> joined at the breakpoint, allow the rate of decline to be different before and after
> menopause while keeping the model interpretable.
>
> **How to interpret:**
>
> - A steeper line = faster rate of hearing decline (more dB lost per year)
> - If the women's line becomes steeper *after* 49.3 yr relative to men's, this supports the
>   hypothesis that menopause accelerates hearing decline in women
> - The scatter around the lines indicates how much individual variation exists beyond the
>   age trend (related to the low R² values seen in Table 3)
 — Piecewise LR scatterplots

*Replicates Fig 3 scattergram.pptx — 6 panels (one per frequency), piecewise regression lines for men and women split at 49.3 yr*

```{r fig3-scatter, fig.height=11, fig.width=9}
# Build predicted lines for each sex × frequency
pred_lines <- map_dfr(c("Men", "Women"), function(sex) {
  map_dfr(FREQS, function(freq) {
    sub <- df %>% filter(Sex == sex)
    mod <- fit_piecewise_ols(sub, freq)

    age_min <- min(sub$Age)
    age_max <- max(sub$Age)

    age_seq <- c(
      seq(age_min, BREAKPOINT, length.out = 60),
      seq(BREAKPOINT, age_max, length.out = 60)
    )

    pred_df <- tibble(
      Age       = age_seq,
      Age_below = pmin(age_seq, BREAKPOINT),
      Age_above = pmax(age_seq - BREAKPOINT, 0)
    )
    pred_df$Predicted <- predict(mod, newdata = pred_df)

    pred_df %>%
      mutate(Sex = sex, Freq = freq, Segment = if_else(Age <= BREAKPOINT, "below", "above"))
  })
}) %>%
  mutate(Freq = factor(Freq, levels = FREQS),
         Sex  = factor(Sex, levels = c("Men", "Women")))

# Raw scatter data
scatter_data <- df %>%
  pivot_longer(all_of(FREQS), names_to = "Freq", values_to = "Threshold") %>%
  mutate(Freq = factor(Freq, levels = FREQS))

# Plot — 3 rows × 2 cols (left col: low freq, right col: high freq)
freq_order <- c("0.25kHz", "1kHz", "4kHz", "0.5kHz", "2kHz", "8kHz")

make_scatter_panel <- function(freq_label) {
  scatter_data %>%
    filter(Freq == freq_label) %>%
    ggplot(aes(x = Age, y = Threshold, color = Sex)) +
    geom_point(alpha = 0.12, size = 0.8) +
    geom_line(data = pred_lines %>% filter(Freq == freq_label),
              aes(x = Age, y = Predicted, color = Sex, group = interaction(Sex, Segment)),
              linewidth = 1.2) +
    geom_vline(xintercept = BREAKPOINT, linetype = "dotted",
               color = "grey40", linewidth = 0.8) +
    scale_color_manual(values = SEX_COLORS) +
    scale_x_continuous(breaks = c(20, 30, 40, 49.3, 60, 70)) +
    labs(title = freq_label, x = "Age (years)", y = "Threshold (dB HL)") +
    theme_hearing() +
    theme(legend.position = "none",
          plot.title = element_text(size = 11))
}

panels <- map(FREQS, make_scatter_panel)

wrap_plots(panels, ncol = 2) +
  plot_annotation(
    title    = "Figure 3. Piecewise Linear Regression: Age vs Hearing Threshold",
    subtitle = paste0("Breakpoint = ", BREAKPOINT, " yr (Korean mean menopause age)  |  Lines = OLS fit per segment"),
    theme    = theme(plot.title    = element_text(face = "bold", size = 13),
                     plot.subtitle = element_text(size = 10, color = "grey40"))
  ) &
  theme(legend.position = "none")

# Shared legend
legend_plot <- ggplot(
  tibble(Sex = factor(c("Men", "Women"), levels = c("Men", "Women")), x = 1, y = 1),
  aes(x, y, color = Sex)) +
  geom_point() +
  scale_color_manual(values = SEX_COLORS) +
  theme_void() +
  theme(legend.position = "bottom",
        legend.title = element_blank())
legend_plot
```

## 6.2 Figure 4 — Rate of decline with bootstrap BCa CI

> **What this figure shows:** The estimated rate of hearing decline (dB/year) for men and women
> at each frequency, separately for the under-49.3 and over-49.3 age groups. Error bars are
> 95% bootstrap confidence intervals.
>
> **How to interpret:**
>
> - **Bar height** = rate of threshold increase per year (higher = faster hearing loss)
> - **Error bars** = uncertainty around that estimate; if bars for men and women do not
>   overlap, the sex difference in rate is likely statistically significant
> - **Comparing the two panels** (below vs. above 49.3): if women's bars grow taller relative
>   to men's in the right panel, this indicates a post-menopausal acceleration
> - **Frequency pattern:** Rates typically increase from low (0.25 kHz) to high (8 kHz)
>   frequencies, reflecting that high-frequency hearing declines fastest with age
 — Rate of decline with bootstrap BCa CI

*Replicates Fig 4.pptx — bar chart of estimated rates (dB/year) with 95% CI error bars*

```{r fig4-rate, fig.height=6, fig.width=11}
rate_data <- map_dfr(c("Men", "Women"), function(sex) {
  map_dfr(FREQS, function(freq) {
    key <- paste(sex, freq, sep = "_")
    r   <- boot_results[[key]]
    bind_rows(
      tibble(Sex = sex, Freq = freq,
             Segment  = paste0("< ", BREAKPOINT, " yr"),
             Slope    = as.numeric(r$slope_below),
             CI_lo    = r$CI_below[1],
             CI_hi    = r$CI_below[2]),
      tibble(Sex = sex, Freq = freq,
             Segment  = paste0(">= ", BREAKPOINT, " yr"),
             Slope    = as.numeric(r$slope_above),
             CI_lo    = r$CI_above[1],
             CI_hi    = r$CI_above[2])
    )
  })
}) %>%
  mutate(
    Freq    = factor(Freq, levels = FREQS),
    Sex     = factor(Sex, levels = c("Men", "Women")),
    Segment = factor(Segment, levels = c(paste0("< ", BREAKPOINT, " yr"),
                                          paste0(">= ", BREAKPOINT, " yr")))
  )

ggplot(rate_data,
       aes(x = Freq, y = Slope, fill = Sex,
           group = interaction(Sex, Segment),
           color = Sex)) +
  geom_col(aes(fill = Sex),
           position = position_dodge(width = 0.7),
           width = 0.6, alpha = 0.85) +
  geom_errorbar(aes(ymin = CI_lo, ymax = CI_hi),
                position = position_dodge(width = 0.7),
                width = 0.25, linewidth = 0.9) +
  geom_hline(yintercept = 0, linewidth = 0.5) +
  facet_wrap(~ Segment) +
  scale_fill_manual(values  = SEX_COLORS) +
  scale_color_manual(values = SEX_COLORS) +
  labs(
    title    = "Figure 4. Estimated Rate of Hearing Decline by Sex and Frequency",
    subtitle = "Bootstrap percentile 95% CI (n = 2,000 resamples)  |  dB/year",
    x        = "Frequency (kHz)",
    y        = "Rate of Decline (dB/year)",
    fill     = NULL, color = NULL
  ) +
  theme_hearing()
```

## 6.3 OLS vs WLS vs Robust — slope comparison

> **Why compare three methods?**
>
> We showed in Phase 4 that heteroscedasticity and non-normality are present. Different
> methods handle this differently:
>
> | Method | What it addresses | Trade-off |
> |---|---|---|
> | **OLS** | Nothing — baseline model | SEs may be too small if heteroscedastic |
> | **WLS** | Heteroscedasticity (gives less weight to high-variance age groups) | Requires estimating weights |
> | **Robust (HC3)** | Heteroscedasticity via sandwich SEs | Slightly more conservative than OLS |
> | **Bootstrap** | Non-normality and heteroscedasticity | Computationally intensive |
>
> **Interpretation:** If all four methods produce similar slope estimates, the results are
> robust — the choice of method does not matter much. If OLS slopes differ substantially from
> WLS/robust slopes, it suggests the heteroscedasticity was meaningfully biasing the OLS
> estimates and the corrected estimates should be reported.
 — slope comparison table

```{r slope-comparison}
comparison_df <- map_dfr(c("Men", "Women"), function(sex) {
  sub <- df %>% filter(Sex == sex)
  map_dfr(FREQS, function(freq) {
    m_ols <- fit_piecewise_ols(sub, freq)
    m_wls <- fit_piecewise_wls(sub, freq)
    m_rob <- fit_piecewise_robust(sub, freq)
    key   <- paste(sex, freq, sep = "_")
    bsr   <- boot_results[[key]]

    map_dfr(list(c("Age_below", "Below 49.3"), c("Age_above", "Above 49.3")),
            function(p) {
      param <- p[1]; seg <- p[2]
      tibble(
        Sex          = sex,
        Freq         = freq,
        Segment      = seg,
        OLS_slope    = round(coef(m_ols)[param], 4),
        OLS_SE       = round(summary(m_ols)$coefficients[param, "Std. Error"], 4),
        WLS_slope    = round(coef(m_wls)[param], 4),
        WLS_SE       = round(summary(m_wls)$coefficients[param, "Std. Error"], 4),
        Robust_slope = round(m_rob[param, "Estimate"], 4),
        Robust_SE    = round(m_rob[param, "Std. Error"], 4),
        Boot_CI_lo   = round(if (seg == "Below 49.3") bsr$CI_below[1] else bsr$CI_above[1], 4),
        Boot_CI_hi   = round(if (seg == "Below 49.3") bsr$CI_below[2] else bsr$CI_above[2], 4)
      )
    })
  })
})

print(comparison_df)
write_csv(comparison_df, "slope_comparison_OLS_WLS_Robust_Bootstrap.csv")
cat("Saved: slope_comparison_OLS_WLS_Robust_Bootstrap.csv\n")
```

## 6.4 Sex-difference interaction test (FDR-corrected)

> **Why this test, not separate regressions?**
>
> The original manuscript ran separate regressions for men and women and then visually
> compared the slopes. The statistically correct way to test whether two slopes are
> *significantly different from each other* is to include both groups in a single model with
> an **interaction term** (Age × Sex). This provides a formal p-value for the sex difference
> in slope.
>
> **How the model works:**
>
> `Threshold = B0 + B1·Age_seg + B2·SexCode + B3·(Age_seg × SexCode)`
>
> - B1 = slope for men (reference group)
> - B1 + B3 = slope for women
> - **B3 (the interaction coefficient)** = difference in slopes (Women minus Men)
> - **p-value of B3** = tests H₀: slopes are equal between men and women
>
> HC3 robust standard errors are used because of the heteroscedasticity identified in Phase 4.
> BH FDR correction is applied because we run this test 12 times (6 frequencies × 2 segments).
>
> **Interpretation:**
>
> - `Interaction_coef > 0` → women's slope is steeper (faster decline) than men's
> - `Interaction_coef < 0` → men's slope is steeper
> - `Sig_FDR = TRUE` → the sex difference in rate is statistically significant after
>   correcting for multiple comparisons
 (FDR-corrected)

```{r interaction-test}
sex_diff_df <- map_dfr(FREQS, function(freq) {
  map_dfr(list(c("below", paste0("< ", BREAKPOINT)),
               c("above", paste0(">= ", BREAKPOINT))), function(grp) {
    seg <- grp[1]; seg_label <- grp[2]

    d <- df %>%
      { if (seg == "below") filter(., Age < BREAKPOINT) else filter(., Age >= BREAKPOINT) } %>%
      select(Age, SexCode, all_of(freq)) %>%
      rename(Threshold = all_of(freq)) %>%
      drop_na() %>%
      mutate(
        Age_seg  = if (seg == "below") Age else Age - BREAKPOINT,
        interact = Age_seg * SexCode
      )

    mod <- lm(Threshold ~ Age_seg + SexCode + interact, data = d)
    rob <- coeftest(mod, vcov = vcovHC(mod, type = "HC3"))

    tibble(
      Freq             = freq,
      Segment          = seg_label,
      Slope_Men        = round(rob["Age_seg",  "Estimate"], 4),
      Slope_Women      = round(rob["Age_seg",  "Estimate"] +
                               rob["interact", "Estimate"], 4),
      Interaction_coef = round(rob["interact", "Estimate"], 4),
      Interaction_p    = round(rob["interact", "Pr(>|t|)"], 4)
    )
  })
})

sex_diff_df <- sex_diff_df %>%
  mutate(
    FDR_p   = round(p.adjust(Interaction_p, method = "BH"), 4),
    Sig_FDR = FDR_p < 0.05
  )

cat("=== Sex Difference in Slopes (Interaction Test, HC3 SE, BH-FDR) ===\n")
print(sex_diff_df)
write_csv(sex_diff_df, "slope_sex_diff_test.csv")
cat("Saved: slope_sex_diff_test.csv\n")
```

## 6.5 Sensitivity analysis — varying the breakpoint

> **Why this matters:** The 49.3-year breakpoint is the *average* Korean menopause age, but
> any individual woman may experience menopause earlier or later. If the study's conclusions
> depend heavily on choosing exactly 49.3, the findings would be fragile. A sensitivity
> analysis tests whether results change meaningfully when the breakpoint is shifted by ±1 or
> ±2 years.
>
> **How to interpret:** If slope estimates are stable across all five breakpoints (47.3, 48.3,
> 49.3, 50.3, 51.3), the findings are robust to the exact menopause age used. If the slopes
> change dramatically, it suggests the results are sensitive to this assumption and should be
> reported more cautiously. Flat lines in the sensitivity plot = robust findings.
 — varying the breakpoint

```{r sensitivity, cache=TRUE}
BREAKPOINTS_TEST <- c(47.3, 48.3, 49.3, 50.3, 51.3)

sens_df <- map_dfr(BREAKPOINTS_TEST, function(bp) {
  map_dfr(c("Men", "Women"), function(sex) {
    sub <- df %>% filter(Sex == sex)
    map_dfr(FREQS, function(freq) {
      d <- make_piecewise_df(sub, freq, bp)
      tryCatch({
        mod <- lm(Threshold ~ Age_below + Age_above, data = d)
        rob <- coeftest(mod, vcov = vcovHC(mod, type = "HC3"))
        tibble(
          Breakpoint   = bp,
          Sex          = sex,
          Freq         = freq,
          Slope_below  = round(rob["Age_below", "Estimate"], 4),
          Slope_above  = round(rob["Age_above", "Estimate"], 4),
          R2           = round(summary(mod)$r.squared, 4)
        )
      }, error = function(e) NULL)
    })
  })
})

cat("=== Slope sensitivity to breakpoint choice ===\n")
sens_df %>%
  group_by(Breakpoint, Sex) %>%
  summarise(across(c(Slope_below, Slope_above, R2), mean), .groups = "drop") %>%
  print()

write_csv(sens_df, "sensitivity_breakpoints.csv")
cat("Saved: sensitivity_breakpoints.csv\n")
```

```{r fig-sensitivity, fig.height=7}
sens_df %>%
  mutate(Freq = factor(Freq, levels = FREQS),
         Sex  = factor(Sex, levels = c("Men", "Women"))) %>%
  pivot_longer(c(Slope_below, Slope_above),
               names_to = "Segment", values_to = "Slope") %>%
  mutate(Segment = recode(Segment,
                          Slope_below = "Below breakpoint",
                          Slope_above = "Above breakpoint")) %>%
  ggplot(aes(x = Breakpoint, y = Slope,
             color = Sex, linetype = Segment, group = interaction(Sex, Segment))) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2.5) +
  geom_vline(xintercept = BREAKPOINT, linetype = "dotted",
             color = "grey40", linewidth = 0.8) +
  facet_wrap(~ Freq, nrow = 2) +
  scale_color_manual(values = SEX_COLORS) +
  labs(title = "Slope Sensitivity to Breakpoint Choice (±2 yr around 49.3)",
       x = "Breakpoint (years)", y = "Slope (dB/year)",
       color = NULL, linetype = NULL) +
  theme_hearing()
```

## 6.6 PSM sensitivity — ANCOVA on matched sample

> **Why repeat the analysis on the matched sample?**
>
> This is the final check that the sex-group imbalance (910 men vs 3,175 women) was not
> inflating or suppressing the observed sex differences. If the ANCOVA results on the matched
> sample (where both groups are age-balanced) agree with the full-sample results, we can
> confidently report that the findings are not an artefact of the unequal group sizes.
>
> **Interpretation:** Compare `FDR_p` values and `SexCoef_dB` between the full-sample ANCOVA
> (Section 5.2) and this matched-sample ANCOVA. Similar coefficients and significance patterns
> = the imbalance was not a problem. Notably different results = the imbalance was influencing
> conclusions and should be explicitly discussed.
 — repeat ANCOVA on matched sample

```{r psm-sensitivity}
ancova_matched <- collect_ancova_pvalues(df_matched) %>%
  mutate(
    FDR_p       = round(p.adjust(Raw_p, method = "BH"), 4),
    Significant = FDR_p < 0.05
  )

cat("=== ANCOVA on PSM-Matched Sample ===\n")
print(ancova_matched)
```

## 6.7 Final summary table

> **How to read this table:**
>
> - **Slope_dB_yr** = estimated rate of hearing threshold increase per year (from piecewise LR).
>   A value of 0.5 means hearing gets 0.5 dB worse for every additional year of age in that
>   segment.
> - **CI_95** = 95% bootstrap confidence interval for the slope. If this interval excludes 0,
>   the rate of decline is statistically significantly different from zero.
> - **Robust_p** = p-value from the HC3 robust regression for the slope. This is more reliable
>   than the standard OLS p-value when heteroscedasticity is present.
>
> This table is the primary numerical result of the re-analysis and can be used directly to
> update Table 3 (dummy variable regression results) in the manuscript.


```{r final-summary}
summary_df <- map_dfr(c("Men", "Women"), function(sex) {
  sub <- df %>% filter(Sex == sex)
  map_dfr(FREQS, function(freq) {
    key   <- paste(sex, freq, sep = "_")
    bsr   <- boot_results[[key]]
    m_rob <- fit_piecewise_robust(sub, freq)

    bind_rows(
      tibble(
        Sex           = sex,
        Frequency     = freq,
        Age_Segment   = paste0("< ", BREAKPOINT, " yr"),
        Slope_dB_yr   = round(as.numeric(bsr$slope_below), 3),
        CI_95         = paste0("[", round(bsr$CI_below[1], 3),
                               ", ", round(bsr$CI_below[2], 3), "]"),
        Robust_p      = round(m_rob["Age_below", "Pr(>|t|)"], 4)
      ),
      tibble(
        Sex           = sex,
        Frequency     = freq,
        Age_Segment   = paste0(">= ", BREAKPOINT, " yr"),
        Slope_dB_yr   = round(as.numeric(bsr$slope_above), 3),
        CI_95         = paste0("[", round(bsr$CI_above[1], 3),
                               ", ", round(bsr$CI_above[2], 3), "]"),
        Robust_p      = round(m_rob["Age_above", "Pr(>|t|)"], 4)
      )
    )
  })
})

cat("=== Final Summary: Piecewise Slopes with Bootstrap 95% CI ===\n")
print(summary_df)
write_xlsx(summary_df, "final_summary_slopes.xlsx")
cat("Saved: final_summary_slopes.xlsx\n")
```

---

# Summary

| Output file | Contents |
|---|---|
| `slope_comparison_OLS_WLS_Robust_Bootstrap.csv` | OLS / WLS / HC3 / Bootstrap slope estimates |
| `slope_sex_diff_test.csv` | Interaction test for sex differences (FDR-adjusted) |
| `sensitivity_breakpoints.csv` | Slopes across 5 alternative breakpoints |
| `final_summary_slopes.xlsx` | Master table: slopes + BCa CI + robust p-values |

### Statistical checklist

- [x] Data integrity validated (shape, columns, nulls, duplicates)
- [x] PSM performed to address sex-group imbalance (sensitivity)
- [x] Normality tested (Shapiro-Wilk / KS)
- [x] Heteroscedasticity tested (Breusch-Pagan) and visualised
- [x] Outliers identified (Cook's D)
- [x] WLS applied for heteroscedasticity
- [x] HC3 robust SEs applied
- [x] Bootstrap percentile 95% CIs (n = 2,000 resamples)
- [x] FDR correction (BH) across all significance tests
- [x] Piecewise LR at fixed breakpoint 49.3 yr
- [x] Dummy-variable interaction test for slope sex differences
- [x] Sensitivity analysis across breakpoints ±2 yr

```{r session-info}
sessionInfo()
```
